Quantifying and Modeling Global Water Accessibility including Water Typology

### 2021-01-14 docs/manuscript/manuscript.tex
updated manuscript

%=== GRAPHICS PATH ===========
%\graphicspath{{./140408-Images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } }
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

@@

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,backend=biber, natbib,maxcitenames=2,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{../References/Water-Accessibility.bib}

@@

\paragraph{}
\citet{onda2012global}discusses the global access to safe water by accounting for water quality and the resulting impact on MDG progress. In 2012, \citet{onda2012global} altered the current Joint monitoring programme (JMP) estimation by including for microbial water quality and sanitary risk utilizing national water quality data. \citep{onda2012global} used principal components analysis to analyze the national environment and development indicators and then created models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it is done manually. How the modes of transportation can affect one's health.
\paragraph{}

@@

\paragraph{}

@@

\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\ 
  & &  \\  
  & &     
\end{tabular}
\end{center}
\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring. Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access. Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.  
\paragraph{}
\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru. This relates and informs one mode of transportation of clean and accessible water. Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc. Overall, the article found that the percent of households with IWS was a few percentage points greater than of the reported utilities. 
\paragraph{}
\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%). \citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions. \citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities. Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation. Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away. Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals. 
\paragraph{}
\citet{goal} explains specific targets that have goals and indicators to track global water accessibility. The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management. They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
% citep (for citations fully enclosed in parentheses)
% cite (for citations without any parentheses)
% citet (for citations where the year only is in parentheses)
%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
%Another example is \citet{fuller2016tracking}.
%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
%\citet{cassivi2018access} discuss time access...
%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
%Other examples: \citep{amit2019measuring,ho2014challenge}
\subsection{Gaps and areas for improvement}
We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.

@@

\hl{1. We could also add the relevant data extracted from DHS showing what countries were also collected.
2. Show numbers because numbers tells great story.}
\paragraph{•}
\hl{ 3. Describe the 11 water accessibility variables, where it came from \citep{price2019difference}

@@

\section*{Methods}
%\hl{Ultimately, there will be a preamble to summarize the contents of this section}
\paragraph{Preamble}
We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.

\subsection*{Cluster analysis}
\hl{You have all you need to write this section. 
You want to describe the algorithms for clustering, the indices for measuring cluster quality, and process you follow to choose the best method/cluster number.}

\hl{ 1. Description of the algorithms implemented: single, average, complete, ward.D, and ward.D2 }

\hl{ 2. Explanation of using NBClust function in R for ward.D2, ward.D, Complete, and K-means; these four methods were elected due their organization of the initial clustering trees.}

\hl{ 3. List of the optimal number of cluster values and the indices number of values for different alogrithm methods for the clustering methods. Description of how to determine the potential best clustering methodology. }

\hl { 4. Clustering optimization data share and explain: Ward.D2 chosen with 4 clustering tree methodology. Share the finished clustering graph for visualization.}

@@

\subsection*{Model results}

>>>>

%=== GRAPHICS PATH ===========
\graphicspath{{./images/}}
% Marginpar width
%Marginpar width
\newcommand{\pts}[1]{\marginpar{ \small\hspace{0pt} \textit{[#1]} } }
\setlength{\marginparwidth}{.5in}
%\reversemarginpar
%\setlength{\marginparsep}{.02in}

@@

\usepackage{rotating}
%% Cite Title
\usepackage[style=authoryear,backend=biber, natbib,maxcitenames=2,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{../Water-Accessibility.bib}

@@

\citet{onda2012global} discusses the global access to safe water by accounting for water quality and the resulting impact
on MDG progress. They altered the current Joint monitoring programme (JMP) estimation by
including for microbial water quality and sanitary risk utilizing national water quality data. \citet{onda2012global}
used principal components analysis to analyze the national environment and development indicators and then created
models. Overall, our analysis can highlight the potential fatigue illness caused by people transporting the water if it
is done manually. How the modes of transportation can affect one's health.

@@

@@

\begin{center}
\begin{tabular}{ c c c }
 short-term & medium term & long-term \\ 
  & &  \\  
  & &     
\end{tabular}
\end{center}
\citet{price2019difference} recommends future research to explore the short-term water access and quality monitoring. Also we should learn to recognize the role of socio-cultural factors that may impact the temporal dynamics of safe water access. Our project can pinpoint additional water accessibility variables that can be used to focus to recognize the time variations as well.  
\paragraph{}
\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru. This relates and informs one mode of transportation of clean and accessible water. Our research can make inferences about if the customers receive clean water elsewhere, such as water tankers, kiosks, vendors, etc. Overall, the article found that the percent of households with IWS was a few percentage points greater than of the reported utilities. 
\paragraph{}
\citet{deshpande2020mapping} discusses their findings of improved overall water availability and piped water accessibility for the globe (40\% to 50.3\%). \citet{deshpande2020mapping} also found that it was the lowest increase in sub-Saharan Africa, where the accessibility was mostly concentrated in urban regions. \citet{deshpande2020mapping} focuses mostly on understanding trends in diarrhea burden, where the needs are greatest for the improvement of access to safe drinking water and sanitation facilities. Our study can benefit this research by implementing the best and effective modes of transportation of clean water to improve safe drinking water and sanitation. Even with the water pipe access increasing, they can't conclude if the people had clean water as they needed right away. Overall, the access to safe drinking water and sanitation improved globally between 2000 and 2017, but dis-proportionality undermines reaching the SDG goals. 
\paragraph{}
\citet{goal} explains specific targets that have goals and indicators to track global water accessibility. The article quantifies and provides Graphs: safe drinking water, safe sanitation, and hygiene, safe sanitation and hygiene, ambient water quality, water use efficiency, levels of freshwater stress, integrated water management, trans-boundary water cooperation, protect and restore water-related ecosystems, water and sanitation support, local participation in sanitation management. They have many ambitious plans to achieve by 2020 and 2030, and our research can find a correlation to why achieving this may be possible.
% citep (for citations fully enclosed in parentheses)
% cite (for citations without any parentheses)
% citet (for citations where the year only is in parentheses)
%%\textit{Discuss the who/what/when/where? (You already have this... just need to reorganize or improve structure)}
%Another example is \citet{fuller2016tracking}.
%\citet{rawas2020comparing} highlights the need for different approaches of intermittent water supplies (IWS) to monitor piped water supply perpetually in Peru.
%\citet{cassivi2018access} discuss time access...
%\hl{Perhaps insert a table summarizing the various dimensions of water accessibility that have been studied}
%Other examples: \citep{amit2019measuring,ho2014challenge}
\subsection{Gaps and areas for improvement}
We are interested in analyzing the impact of transportation modes (usage/ownership) on water accessibility.

@@

@@

\section*{Methods}

\paragraph{Preamble}
We analyzed data from the Demographic Health Survey (DHS) and the Multiple Indicator Cluster Survey (MICS) (1990-2020) to highlight water accessibility predictors in 78 countries throughout the world.

\subsection*{Cluster analysis}

@@

\begin{figure}[h!]
  \centering
  %\includegraphics[width=.7\textwidth]{spider-plot}
  \caption{Spider plots of water accessibility variables by typology}
  \label{fig:spider}
\end{figure}
\subsection*{Model results}
